Learning by Teaching a Robot:\\The Case of Handwriting

### 2015-07-10 main.tex
Proof-reading second half, except case studies

\title{Learning by Teaching a Robot: the Case of Handwriting}

@@

\resizebox{0.6\linewidth}{!}{%

@@

    \path (nao) edge [->,redarrow, bend left] node[label, auto] {robot state} (laptop);
    \path (laptop) edge [->,greenarrow, bend left] node[label, auto] {writing gesture} (nao);

    \path (tablet) edge [->,redarrow, bend left] node[label, auto] {demonstrations,\\turn taking} (laptop);
    \path (laptop) edge [->,redarrow, bend left] node[label, auto=right] {letter to write} (tablet);

    \path (selection) edge [->,redarrow] node[label, auto] {letter/word to write} (laptop);

@@

\caption{\small Overview of the system. In total, the system runs about 10 ROS nodes,
    distributed over the robot itself, a central laptop and Android tablets.}

@@

The overall architecture of the system (Figure~\ref{fig:archi}) is therefore
spread over several devices: the {\sc nao} robot itself, that we address via
both a ROS API\footnote{The ROS stack for {\sc nao} is available at
\url{http://wiki.ros.org/nao_robot}.} and the Aldebaran-provided NaoQI API, one
to four Android tablets (the main tablet is used to print the robot's letter and
to acquire the children's demonstrations. More tablets have been used in some
studies, either to let the child input words to be written, or for the
experimenter to qualitatively annotate the interaction in a synchronized
fashion), and a central laptop running the machine learning algorithms, the
robot's handwriting gesture generation and high level control of the activity.

Since the system does not actually require any CPU-intensive process, the laptop
could be removed and the whole logic could run on the robot. Due to the relative
difficulty to deploy and debug ROS nodes directly on the robot, we however
effectively kept the laptop during our experiments.

@@

\subsubsection*{Trajectory Following Movements}

Using simulated handwriting provides an opportunity for the robot's writing to
appear smoother than would be achievable with a writing instrument. However, the
robot's motions must still sufficiently match the displayed trajectory in order
capture the engagement of the child participant in the action. Aldebaran's NaoQi
API is used for the inverse kinematics of the trajectory following. ROS is used
for integration of {\sc nao} with external reference frames, such as the
tablet's location, using the $tf$ transformation library~\cite{Foote2013}.

When using simulated handwriting, it is no longer necessary that the robot
engages in the typical style of handwriting of using a writing instrument at a
desk. Having instead the robot point at the writing surface to cause the
trajectory to appear (as seen in Figure \ref{fig:schools}, left) has several advantages:

\begin{itemize}

    \item The working space of the robot increases, both in the technical sense
        and the interaction sense: someone can, in theory, show the tablet to
        the robot from across the room and have it still respond, without
        needing the tablet to be within arm's reach,

    \item Concerns about whether or not the child would start mimicking the
        robot's incorrect writing form (\eg pen grip) are mitigated,

    \item Perhaps most significantly, the accuracy of the matching of the
        robot's motion with the trajectory displayed on the tablet is not as
        critical. This is because a pen tip would be expected to touch the
        tablet exactly at the trajectory point, while a fingertip may not.

\end{itemize}

We have therefore designed the system in such a way that the robot is simulating
handwriting by pointing at the tablet\footnote{Teachers interviewed for their
feedback on the system advised that children are asked to draw letters in
the air in a similar manner as part of their handwriting education. The
behaviour is hence not unfamiliar to children.}. As interacting with a
tablet with one's finger is not uncommon, this may aid the acceptance of the
writing style by users.

Because motion planning is performed with respect to the hand of the robot,
rather than its fingertip, one or two of the orientation degrees of freedom of
the hand are fixed to keep the finger approximately perpendicular to the writing
surface, depending on the desired accuracy. The remaining free orientation(s),
coupled with the whole-body motion control available, allow for a sufficient
working space for writing on the entire tablet.

@@

\begin{table}[ht!]
\centering
\caption{\small Field studies conducted within the project}
\label{studies}
\footnotesize
\begin{tabular}{@{}lp{4cm}p{2.2cm}p{1.8cm}l@{}}
\toprule
{\bf Study}     & {\bf Type}                                    & {\bf Duration}              & {\bf Children \#}                     & {\bf Ages} \\ \midrule
{\it ISG 1}     & Unstructured group interaction at school      & 16 min/group               & 4 $\times$ 8 children                 & 6-7        \\
{\it Florimont} & Individual/Pair interaction at school         & 11 min/group               & 7 (individual) + 7 $\times$ 2 (pairs) & 7-8        \\
{\it Vincent}     & Case-study, spanning over 4 weeks              & 1.5 hours $\times$ 4 weeks & 1                                     & 5          \\
{\it Haut-Lac}  & Pair (with turn-taking) interaction at school & 26 min/group                            & 7 $\times$ 2                          & 5-6        \\
{\it ISG 2}     & Individual interaction at school              & 18 min                            & 6                                     & 5-6        \\
{\it Thomas}     & Case-study, spanning over 4 weeks              & 1h $\times$ 4 weeks        & 1                                     & 5          \\ \bottomrule
\end{tabular}
\end{table}

To date, the presented system has been deployed and tested in three different
schools (totalling child-robot interactions with more than 70 children, aged 5
to 8) and we also conducted 2 case studies over several weeks.
Table~\ref{studies} gives an overview of these studies.

@@

Over the two years of the project, we conducted four studies in schools
(Figure~\ref{fig:schools}). These experiments were meant to technically validate
the system (is it actually able to autonomously write and learn from
demonstrations?) and the interaction (is the apparatus easy to grasp and to
interact with for children), and to study the initial acceptance of the robot in
the school environment (we conducted several formal and informal discussions
with teachers).

We also relied on these four studies to acquire a dataset of children's
handwritten letters and to study how they engage with the robot (and maintain or
not this engagement).

@@

Critically, these studies were conducted with \emph{whole classes}: we did not
select underperforming children that requires careful beforehand organization
with the school to limit ethical issues. Having more children (73 in total) was
however beneficial for these preliminary studies.

@@

From a technical perspective, the system achieves an acceptable level of
reliability and allow a technically sound autonomous interaction. For instance,
during the second school study (\textit{Florimont}), the robot withstood
interactions which lasted for a total of 160 minutes.  During this time the
robot wrote 335 letters, 152 of which in response to demonstrations received
from the 21 children. Technical intervention was only required for the three
instances that the robot fell later in that day.  Otherwise, the technical
components of the system operated autonomously and as expected over the
sessions.

Due to the modular software architecture (about 10 ROS nodes), crashes occurring
during the sessions were usually quickly resolved by re-launching the faulty
node alone, and did not significantly impact the interaction.

The otherwise technical limitations were related to some letters or writing
style (most notably, the ones requiring multiple strokes per letter) not being
adequately processed by the learning algorithm. Support for such letters has
been since added.

@@

For the approach to be effective, it is critical that children recognize and
accept that the robot is writing by itself. When asked, no child indicated that
they did not believe that the robot was writing by itself. There were, at times,
questions about the robot's writing method at the beginning of the interaction,
but when advised that the robot ``tells the tablet what it wants to write,''
this was accepted by the children.  Besides, teachers interviewed for their
feedback on the system advised that children are asked to draw letters in the
air in a similar manner as part of their handwriting education. The behaviour is
hence not unfamiliar to children.

\paragraph{Engagement}

As indicated in the introduction, literature~\cite{Hoy2011} suggests that a
total of 20 handwriting practice sessions is found to be the minimum to demonstrate effective
results in handwriting remediation. This highlights the necessity to sustain
engagement over the long-term if we want to achieve measurable learning gains.

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.9\linewidth]{engagement}
    \caption{Engagement level of one child, manually annotated by two judges
        over a 13 min long interaction (Cronbach’s alpha test: $\alpha = .82$,
        good inter-judge agreement). 0 means ``complete disengagement from the
        task'', 6 means ``full engagement''. The vertical lines represent activity
    switches between handwriting teaching and story-telling (used as a
    distruber).}
    \label{engagement_level}
\end{figure}

We also conducted a qualitative assessment of the engagement levels during the
interaction sessions (studies \textit{Haut-Lac} and \textit{ISG 2}).
Figure~\ref{engagement_level} shows the result for one child, with a moderate
level of engagement. On-going research in our group aims at assessing such
engagement levels in real-time, and adapt accordingly its behaviour.

Other qualitative assessment of engagement over longer interactions were
conducted during the case studies, as reported in the following section.

@@

    \caption{\small Text generated by the robot, before and after training with
    the child. As an example, the red box emphasizes the changes on the word
    ``envoyer''.}
    \label{fig:stimuli}
\end{figure}

@@

The CoWriter project, as presented above, gives a novel perspective on educative
robots, at several levels:

@@

    \item this proves essential to \textbf{sustain a long-term interaction} (several
        hours) around a fundamentally routine -- yet challenging -- educational
        task (hand-writing learning).
\end{itemize}

No surprisingly, the role of the robot \textit{vis-a-vis} the teacher has also
been questioned (including an angry email from one teacher: ``You want to
replace us!''): as we see it, the role of the robot within the classroom (or at
the therapist's surgery) does not infringe upon the role of the adult (teacher
or therapist).  The core of the \emph{learning by teaching} paradigm relies on
the child becoming the teacher of an underperforming pupil (the robot): from
that perspective, the robot does not replace the teacher, on the contrary. It
plays a different role in the classroom, which happen to be novel as well: the
robot is the \emph{least} performing student, and however a very patient, always
eager to improve, one.

The importance of the adult is further supported by our experiments: even with
an autonomous, nominally performing robot, to put the teacher's clothes on
remains (expectedly) difficult for 5-6 years old children, and during the
experiments we conducted, the adult always played a key role at prompting the
child to give feedback to the robot or to go to the next letter or word.  At a
higher orchestration level (and as reported in the two case studies with Vincent
and Thomas), the educational scenarios were also always designed and monitored by
the adults.

We initially envisioned the CoWriter system to be run in the back of a
classroom with one child: this would have allowed an individual, face-to-face
remediation approach, not otherwise tractable for a teacher with 20 pupils.
This is unlikely to happen. In our experience, the teacher keeps such an
important role that the interaction and the learning would hardly occur if the
child is left alone (or even semi-alone). At the end, we can reassure teachers:
robots are not going to replace them any time soon.

Beyond handwriting, we believe that this work provides a novel perspective on
the role for robots in the field of education. \emph{Learning by teaching} is a
powerful paradigm because of not only its pedagogical efficacy, but its
potential to positively impact the child's motivation and self-esteem. We hope that 
this article shows that this is a very relevant context of use for robots. Indeed,
when facing a child with school difficulties, robots can play the role of a na\"ive 
learner which neither adults nor peers -- because of the social effects it would 
induce -- can convincingly play.

>>>>

\title{Learning by Teaching a Robot:\\The Case of Handwriting}

@@

\resizebox{0.7\linewidth}{!}{%

@@

    \node[draw,above right=2 of laptop,anchor=north west,text width=4cm] (processes)
    {\sf\scriptsize machine-learning, \\letters/gestures
    generation, \\interaction supervision};
    \path (laptop) edge[dashed] (processes);

    \path (nao) edge [->,redarrow, bend left] node[label, auto] {robot state} (laptop);
    \path (laptop) edge [->,greenarrow, bend left] node[label, auto] {writing gestures} (nao);

    \path (tablet) edge [->,redarrow, bend left] node[label, auto] {demonstrations,\\turn taking} (laptop);
    \path (laptop) edge [->,redarrow, bend left] node[label,
    auto=right,align=right]
    {path of\\ letters to display} (tablet);

    \path (selection) edge [->,redarrow] node[label, auto=right] {letter/word to write} (laptop);

@@

\caption{\small \textbf{Overview of the system}. In total, the system runs about 10 ROS nodes,
    distributed over the robot itself, a central laptop and Android tablets.}

@@

The overall architecture of the system (Figure~\ref{fig:archi}) is therefore
spread over several devices: the {\sc nao} robot itself, that we address via
both a ROS API\footnote{The ROS stack for {\sc nao} is available at
\url{http://wiki.ros.org/nao_robot}.} and the Aldebaran-provided NaoQI API, one
to four Android tablets (the main tablet is used to print the robot's letter and
to acquire the children's demonstrations; more tablets have been used in some
studies, either to let the child input words to be written, or for the
experimenter to qualitatively annotate the interaction in a synchronized
fashion), and a central laptop running the machine learning algorithms, the
robot's handwriting gesture generation and high level control of the activity.

Since the system does not actually require any CPU-intensive process, the laptop
can be removed and the whole logic run on the robot. Due to the relative
difficulty to deploy and debug ROS nodes directly on the robot, the laptop
remains however convenient during the development phase and we kept if during
our experiments.

@@

@@

To date, the system has been deployed and tested in three different
schools (totalling child-robot interactions with more than 70 children, aged 5
to 8) and we also conducted 2 case studies that lasted several weeks.
Table~\ref{studies} gives an overview of these studies.

\begin{table}[ht!]
\centering
\caption{\small Field studies conducted within the project}
\label{studies}
\footnotesize
\begin{tabular}{@{}lp{4cm}p{2.2cm}p{2cm}l@{}}
\toprule
{\bf Study}     & {\bf Type}                                    & {\bf Avg. duration}              & {\bf Children \#}                     & {\bf Ages} \\ \midrule
{\it ISG 1}     & Unstructured group interaction at school      & 16 min/group               & 4 $\times$ 8 children                 & 6-7        \\
{\it Florimont} & Individual/Pair interaction at school         & 11 min/group               & 7 (individual) + 7 $\times$ 2 (pairs) & 7-8        \\
{\it Vincent}     & Case-study, spanning over 4 weeks              & 1.5h $\times$ 4 weeks & 1                                     & 6          \\
{\it Haut-Lac}  & Pair interaction at school & 26 min/group                            & 7 $\times$ 2                          & 5-6        \\
{\it ISG 2}     & Individual interaction at school              & 18 min                            & 6                                     & 5-6        \\
{\it Thomas}     & Case-study, spanning over 4 weeks              & 1h $\times$ 4 weeks        & 1                                     & 5          \\ \bottomrule
\end{tabular}
\end{table}

@@

Over the two years of the project, we conducted four studies in schools
(Figure~\ref{fig:schools}). These experiments were meant to technically validate
the system (is it actually able to autonomously write and learn from
demonstrations?) and test the interaction (is the apparatus easy to grasp and to
interact with for children?). We also studied the initial acceptance of the
robot in the school environment (through several formal and informal discussions
with teachers) and how children engage with the robot (and maintain or not this
engagement).

@@

Critically, these studies were conducted with \emph{whole classes}: we decided
not to select specifically underperforming children as having more children (73
in total) was beneficial for these preliminary studies, and this would have
besides required careful organization with the school due to ethical concerns.

@@

From a technical perspective, the system achieves an acceptable level of
reliability and allow a technically sound autonomous interaction. For instance,
during the second school study (\textit{Florimont}), the robot withstood
interactions which lasted for a total of 160 minutes.  During this time the
robot wrote 335 letters, 152 of which in response to demonstrations received
from the 21 children. Technical intervention was only required for the three
instances that the robot fell during that day.  Otherwise, the technical
components of the system operated autonomously and as expected over the
sessions.

Due to the modular software architecture (about 10 ROS nodes), the occasional
crashes occurring during others sessions were usually quickly resolved by
re-launching the faulty node alone, and did not significantly impact the
interaction.

The otherwise technical limitations were related to some letters or writing
style (most notably, the ones requiring multiple strokes per letter) not being
adequately processed by the learning algorithm. Support for such letters has
been added as a follow-up of the studies.

@@

Children's recognition that the robot is writing by itself is critical for our
approach to be effective. When asked, no child indicated that they did not
believe that the robot was writing by itself. There were, at times, questions
about the robot's writing method at the beginning of the interaction, but when
advised that the robot ``tells the tablet what it wants to write,'' this was
accepted by the children.  Besides, teachers interviewed for their feedback on
the system advised that children are asked to draw letters in the air in a
similar manner as part of their handwriting education. The behaviour is hence
not unfamiliar to children.

More generally, the approach was well accepted and recognized as useful and
promising by teachers and parents; this was \textit{a posteriori} confirmed by
the numerous spontaneous contacts made by parents and therapists who were
looking forward to use the system with their children.

\paragraph{Sustained engagement}

As indicated in the introduction, the literature suggests that 20 handwriting
practice sessions is found to be the minimum to demonstrate effective results in
handwriting remediation~\cite{Hoy2011}. This highlights the necessity to sustain
a child-robot engagement over the long-term if we want to achieve measurable learning gains.

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.9\linewidth]{engagement}
    \caption{\small Engagement level of one child, manually annotated by two judges
        over a 13 min long interaction (Cronbach’s alpha test: $\alpha = .82$,
        good inter-judge agreement). 0 means ``complete disengagement from the
        task'', 6 means ``full engagement''. The vertical lines represent activity
    switches between handwriting teaching and story-telling (used as a
    distruber).}
    \label{engagement_level}
\end{figure}

We also conducted a qualitative assessment of the engagement levels during
several interaction sessions (studies \textit{Haut-Lac} and \textit{ISG 2}).
While the overall engagement was high, Figure~\ref{engagement_level} shows an
example of a child with moderate level of engagement. On-going research in our
group aims at assessing such engagement levels in real-time, so that the robot
can address possible disengagement (by changing/adapting its physical behaviour,
by switching to different activities, etc.)

Further qualitative assessment of engagement over longer interactions were
conducted during the case studies, as reported in the following section.

@@

    \caption{\small Text generated by the robot, before and after training with
        the child (in French). As an example, the red box emphasizes the changes on the word
    ``envoyer''.}
    \label{fig:stimuli}
\end{figure}

@@

We believe that this research provides a novel perspective on educative
robots at several levels. We show that:

@@

    \item this proves essential to \textbf{sustain a long-term interaction} (several
        hours) around a fundamentally routine -- yet challenging -- educational
        task (hand-writing learning),

    \item besides, the social commitment induces a \textbf{cognitive engagement}
        of the child with the robot which is \textbf{a key learning lever} as
        its elicits \textbf{reflective, meta-cognitive mechanisms on the
        learning task}.
\end{itemize}

The place and role of the robot \textit{vis-\`a-vis} the teacher can be
questioned: as we see it, the role of the robot within the classroom (or at the
therapist's surgery) does not infringe upon the role of the adult (teacher or
therapist).  The core of the \emph{learning by teaching} paradigm relies on the
child becoming the teacher of an underperforming pupil (the robot): from that
perspective, the robot does not replace the teacher, on the contrary. It plays a
different role in the classroom, which happen to be novel as well: the robot is
the \emph{least} performing student, and still a very patient, always eager to
improve, one.

The importance of the adult is further supported by our experiments: even with
an autonomous, nominally performing robot, to put the teacher's clothes on
remains (expectedly) difficult for 5-6 years old children, and during the
experiments we conducted, the adult always played a key role at prompting the
child to give feedback to the robot or to move to the next letter or word.  At a
higher orchestration level (and as reported in the two case studies with Vincent
and Thomas), the educational scenarios were also always designed and monitored
by the adults.

We initially envisioned our system to be run in the back of a classroom with one
child: this would have allowed an individual, face-to-face remediation approach,
not otherwise tractable for a teacher with 20 pupils.  This is however unlikely
to happen soon. In our experience, the teacher keeps such an important role that
the interaction (and the learning!) would hardly occur if the child is left
alone (or even semi-alone). We can therefore reassure teachers: robots are not
going to replace them any time soon.

Beyond handwriting, we however do believe that this work provides a novel
perspective on the role for robots in the field of education. \emph{Learning by
teaching} is a powerful paradigm because of not only its pedagogical efficacy,
but its potential to positively impact the child's motivation and self-esteem.
We hope that this article shows that this is a very relevant context of use for
robots: when facing a child with school difficulties, robots can play the role
of a na\"ive learner which neither adults nor peers -- because of the social
effects it would induce -- can convincingly play.
